Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\

### 2025-01-14 main-disk-model.tex
put first figures into the text, described the setups. Everything WIP.

    \subsubsection{Target computing systems}
%    \skug{Because of the normalized nature of the task/edge weights, I will now be saying that our larger cluster has 192 GB and it
%    is represented as 192 000 (with 3 zeroes). For the memory-constrained cluster (192 00 with 2 zeroes) I will say that it has 10 times
%    less memory - 19.2 GB. Other processors will be dealt with similarly.}

    To fully benefit from the historical data, the  {\em default} experimental environment
    that we consider is a cluster based on the same six
    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
    We set the number of each kind of node to $12$, thus having 72 processors in total.  % the whole cluster.
%
    Each machine has a (normalized) CPU speed and a memory size (in GB), and we list them as (name, speed, memory):
    ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines;
    ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and
    large memory (see Table~\ref{tab:procs}).

    \begin{table}[htb]
        \begin{center}
            \begin{tabular}{c|c|cc}
                \toprule
                Processor  %& Amount
                &  CPU speed   & \multicolumn{2}{c}{Memory size (GB)} \\
                name & (GHz) & {\em default} & {\em mem-constrained} \\
                \midrule
                local                    & 4                    & 16     & 1.6 \\
                A1                      & 32                   & 32     & 3.2 \\
                A2                      & 6                    & 64    & 6.4 \\
                N1                      & 12                   & 16     & 1.6 \\
                N2                      & 8                    & 8      & 0.8\\
                C2                      & 32                   & 192   &  19.2\\
                \bottomrule
            \end{tabular}
        \end{center}
        \caption{Cluster configuration.}
        \label{tab:procs}
    \end{table}

    We also consider a more constrained setting, by varying the cluster configuration. The
     {\em memory-constrained cluster} is consisting of 72 nodes (12 of each kind) as the default cluster,
     but each node has 10 times less memory. Hence,  
     the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB, 
     $A2$ has $6.4$~GB instead, of $64$~GB etc.
    The processor speeds and their relations stay unchanged (see Table~\ref{tab:procs}).

Note that in both clusters, we set the size of the communication buffer to be equal
to ten times the the memory size.

@@

    To simulate the execution of a workflow, we implemented a runtime system.
    It reads the historical data and builds weights for tasks, as explained above.
%    These weights are then treated as the ``truth'' about the execution during the simulation.
    In the static case, these values are being sent to the scheduler, which builds a schedule
    according to these weights.
    However, in the dynamic setting, the runtime system applies a deviation function to the values.
    The deviation function computes a normally distributed deviated value, where the initial value
    is the mean and the deviation is~$10\%$.
    This scenario corresponds to the real-life scenarios identified in~\cite{lotaru} and
    other works dedicated to predicting runtimes of tasks~\cite{da2015online,da2013toward}.

    Hence, the scheduler receives deviated values and makes decision based on them.
    This leads to several types of possible issues:
    \begin{itemize}
        \item A processor is blocked by another task. If the scheduler underestimated the execution time of a task,
        it will block another one from starting.
        \item A predecessor has not yet finished.
        The scheduler may request a task to start on its processor, while
        some of the predecessors of the tasks have in fact not yet completed their execution, and the
        task is therefore not yet ready.
%        A similar situation, but a task could start on its processor, but one or more of its parents a
%        are not ready yet.
        \item Not enough memory. If the scheduler underestimated the amount of memory a task requires, this task might not be able to execute on a chosen processor.
        \item A task took less time than expected. We only consider this case if a task took more than 10\% less
        time than expected. In this
        case, we want to exploit the newly acquired free time by possibly starting other tasks earlier.
    \end{itemize}

@@

   % \subsubsection{Preliminary experiments}
We first study the heuristics behavior when the weights do not change upon runtime,
hence the scheduler has a perfect knowledge of task memory requirements and execution times.
We have compared two eviction strategies, starting with large files first or small files first,
and did not report any significant changes in terms of validity of schedule or makespan.
Hence, we only present results with the eviction of largest files first.
%    Our preliminary experiments have shown that evicting large or small files does not give a notable change in makespan.

    \subsubsection{Scheduling on the default cluster}

%    \paragraph{Normal cluster}

>>>>

    \subsubsection{Target computing systems}

    To fully benefit from the historical data, the  {\em default} experimental environment
    that we consider is a cluster based on the same six
    kinds of real-world machines that were used in the experimental evaluation in~\cite{lotaru}.
    We set the number of each kind of node to $6$, thus having 36 processors in total.  % the whole cluster.
%
    Each machine has a normalized CPU speed and a normalized memory size, and we list them as (name, speed, memory):
    ($local$, 4, 16) -- very slow machines; ($A1$, 32, 32), ($A2$, 6, 64), ($N1$, 12, 16) -- average machines;
    ($N2$, 8, 8) -- machine with very small memory; and ($C2$, 32, 192) -- {\em luxury} machine with high speed and
    large memory (see Table~\ref{tab:procs}).

    \begin{table}[htb]
        \begin{center}
            \begin{tabular}{c|c|cc}
                \toprule
                Processor  %& Amount
                &  CPU speed   & \multicolumn{2}{c}{Memory size (relative)} \\
                name & (relative) & {\em default} & {\em mem-constrained} \\
                \midrule
                local                    & 4                    & 16     & 1.6 \\
                A1                      & 32                   & 32     & 3.2 \\
                A2                      & 6                    & 64    & 6.4 \\
                N1                      & 12                   & 16     & 1.6 \\
                N2                      & 8                    & 8      & 0.8\\
                C2                      & 32                   & 192   &  19.2\\
                \bottomrule
            \end{tabular}
        \end{center}
        \caption{Cluster configuration.}
        \label{tab:procs}
    \end{table}

    When we consider a default execution environment, we multiply the relative memory size with $10^6$.
    We multiply the relative speeds with $100$, and let read-write operations unpenalized (multiple the size of files with 1).
    Effectively, we thus reward the computations with $100$ in comparison  with IO operations.
    We further penalize offloading of memory by multiplying its execution time by $1000$.

    For a memory-constrained scenario, we give the processors ten time less memory: we multiply the relative memory size
    with $10^5$.
    We keep all the settings related to speeds the same.
    For a more memory-loo scenario, we give the processors ten times more memory.

    For the IO-based scenarios, we keep the amount of memory the same as in the default scenario and change only the penalties for the read-write
    operations and offloading.
    For the ``IO slower'' case, we penalize both the read-write operations and the offloading ten times more: we multiply the length of
    read-write operations by $10$, and the offloading by $10^4$.
    In the ``IO faster'' case, we do the opposite and reward read-write operations by dividing their time by to, while penalize the offloading only
    by $100$ instead of $1000$.
    Finally, we examine the effect of the offloading penalty reducing and increasing only it ten times for the experimental setups ``offload faster''
    and ``offload slower''.

    TODO: describe with more and less nodes.
   % We also consider a more constrained setting, by varying the cluster configuration. The
   %  {\em memory-constrained cluster} is consisting of 72 nodes (12 of each kind) as the default cluster,
   %  but each node has 10 times less memory. Hence,
   %  the {\em luxury} machine $C2$ has $19.2$~GB memory in this setting instead of $192$~GB,
   %%  $A2$ has $6.4$~GB instead, of $64$~GB etc.
   % The processor speeds and their relations stay unchanged (see Table~\ref{tab:procs}).

@@

    No ``runtime system'' as such in the static case.
    Todo: either remove entirely or explain how simulation of the runtime system works in the dynamic case.

@@

    We first study the heuristics behavior when the weights do not change upon runtime,
    hence the scheduler has a perfect knowledge of task memory requirements and execution times.
    We have compared two eviction strategies, starting with large files first or small files first,
    and did not report any significant changes in terms of validity of schedule or makespan.
    Hence, we only present results with the eviction of largest files first.

    \subsubsection{Relative makespan}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.9\columnwidth] {images/disk-model/rel-ms-normal-bar}
        \caption{Relative makespans on the default cluster setup. Lower is better.}
        \label{fig:rel-ms-normal-bare}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.9\columnwidth] {images/disk-model/rel-ms-iofast-bar}
        \caption{Relative makespans on the cluster setup with faster IO. Lower is better.}
        \label{fig:rel-ms-iofast-bar}
        \vspace{-0.3cm}
    \end{figure}

    \begin{figure}[tb]
        \centering
        \includegraphics[width=0.9\columnwidth] {images/disk-model/rel-ms-ioslow-bar}
        \caption{Relative makespans on the cluster setup with slower IO. Lower is better.}
        \label{fig:rel-ms-ioslow-bar}
        \vspace{-0.3cm}
    \end{figure}

    \subsubsection{Scheduling on the default cluster}
    AFTER HERE IS OLD.
%    \paragraph{Normal cluster}
